Towards a miniKanren with fair search strategies

### 2019-05-24 Paper/paper.tex
some fixes

\usepackage{listings}
\lstset{
  language=Scheme,
  basicstyle=\ttfamily,
  morekeywords={run,conde,run*,defrel,==,fresh,disj,conj,disj+},
  alsodigit={!\$\%&*+-./:<=>?@^_~},
  morecomment=[l]{;,\#{}lang},
  mathescape=true,
  literate={-inf}{{$^\infty$}}{1} {/fair}{{$_{fair}$}}{3}
  		   {→}{{$\rightarrow$}}{1}
  		   {×}{{$\times$}}{1}
  		   {==}{{$\equiv$}}{1}
  		   {conde}{{\textcolor[rgb]{0,.3,.7}{\textbf{cond$^e$}}}}{4}
                   {appendo}{{append$^o$}}{6}
  		   {repeato}{{repeat$^o$}}{6}
  		   {disj2}{{disj$_2$}}{4}
  		   {conj2}{{conj$_2$}}{4}
  		   {g0}{{g$_0$}}{2}
  		   {g1}{{g$_1$}}{2}
  		   {g2}{{g$_2$}}{2}
  		   {gl}{{g$_l$}}{2}
  		   {gr}{{g$_r$}}{2}
  		   {DFSi }{{DFS$_\textrm{i}$}}{3}
  		   {DFSbi}{{DFS$_\textrm{bi}$}}{3}
  		   {DFSf }{{DFS$_\textrm{f}$}}{3}
}
%% scheme-list .

@@

\newcommand{\DFSi }[0]{DFS$_\textrm{i}$}
\newcommand{\DFSf }[0]{DFS$_\textrm{f}$}
\newcommand{\DFSbi}[0]{DFS$_\textrm{bi}$}
\newcommand{\BFS}[0]{BFS}
\newcommand{\BFSser}[0]{BFS$_\textrm{}$}
\newcommand{\BFSimp}[0]{BFS$_\textrm{imp}$}

@@

miniKanren is a family of relational programming languages.
\citep{Friedman:2005:RS:1121583} \citet{friedman_reasoned_2018} introduce 
miniKanren and its implementation in \emph{The Reasoned Schemer} 
and \emph{The Reasoned Schemer, 2nd Ed} (TRS2). \citet{byrd2017unified} have 
demonstrated that miniKanren programs are useful in solving several difficult
problems. miniKanren.org contains the seeds of many difficult problems and 
their solutions.

\NOTE[DPF]{[2,3] is the way to cite this, but I leave cleaning up this
paragraph to you.  Mostly the point is to also include a reference to
the two dimitris and to the first OCamren (OCanren), whatever. It is up to you
in this paragraph.  You can refer to
the two books like this ``Friedman et al[2,3] ... .'' This way the 
paragraph will have only one ``Friedman''}

\NOTE[DPF]{I have decided to agree with you that we should not use 
BFSser, so I am going to remove all occurrences.  
Also, we should say, ``Our implementation of BFS''
and then say, shortened to ``Our BFS'' and then you can use it in the table,
so one column should be Our BFS and the other should be BFS.}
\NOTE[DPF]{Occurrences of BFS (without Our) mean Seres et al. (We need only say
that once, or very occasionally}

\NOTE[DPF]{I think that subscripts i, bi, and f should
match what is in the comments above code, so it should be italic, emph,
or math, whichever was used before.}

\NOTE[DPF]{Is the spelling of reverso truly written without the last e?
Shouldn't it be reverseo with the appropriate raising of o.}

@@

\NOTE[LKC]{The following paragraph is extended to include an explanation on 
\texttt{run*}}

@@

\NOTE[DPF]{Would the answer be the same if the list had been this ((a) (b))
or did you mean structurally the same?}
\NOTE[LKC]{I have rewritten the relevant sentences.}

@@

In this section, we review the implementation of interleaving depth-first 
search (\DFSi). We focus on parts that are relevant to this paper. TRS2,
chapter 10 and the appendix, ``Connecting the wires'', 
provides a comprehensive description of the 
miniKanren 
implementation but limited to unification constraints ($\equiv$).
Fig.~\ref{DFSi-0} and Fig.~\ref{DFSi-1} show parts that are later compared 
with other search 
strategies. We follow some conventions to name variables: \texttt{s}s name 
states; \texttt{g}s (possibly with subscript) name goals; 
variables ending with $^\infty$ name search spaces. Fig.~\ref{DFSi-0} shows the 
implementation of \disj. The 
first function, \disjtwo, implements binary disjunction. It applies the 
two disjunctive goals to the input state \texttt{s} and composes the two 
resulting search spaces with \texttt{append$^\infty$}. The following syntax 
definitions say \disj{} is right-associative. Fig.~\ref{DFSi-1} 
shows the implementation of \conj. The first function, \conjtwo{}, 
implements binary conjunction. 
It applies the \emph{first} goal to the input state, then applies the second 
goal to states in the resulting search space. The helper function 
\texttt{append-map$^\infty$} applies its input goal to states 
in its input search spaces and composes the resulting search spaces. It reuses 
\texttt{append$^\infty$} for search space composition. The following syntax 
definitions say \conj{} is also right-associative.
% \begin{figure}
% 	\lstinputlisting{Figures/conde.rkt}
% 	\caption{implementation of \conde}
% 	\label{conde}
% \end{figure}
% 
% The definition of \conde{} (Fig.~\ref{conde}) is shared by all search 
% strategies. It relates clauses disjunctively, and goals in the same 
% clause conjunctively.

\NOTE[DPF]{sure, go ahead and do it.}
\NOTE[LKC]{I think we can simply remove \conde{} code and its accompany 
description.}

@@

\NOTE[DPF]{I am not in love with the two calls to null?. I can think of at least
two ways to fix it.  Pick one.}

@@

\NOTE[DPF]{The use of the references to Fig 7 twice tells me that something
has to change for the sake of clarity.}

@@

\NOTE[LKC]{I swap \texttt{ks} and \texttt{kf} to mimic \texttt{ind-List}}

\NOTE[DPF]{If we forget to use {} after \reverso, the o ends up too near
the next word.  I fixed reverso on the second line of section 7, but there 
may be others.}

@@

In this subsection, we compare the pure subset of our BFS with \BFS{}. We 
focus on the pure subset because \BFS{} is designed for a pure 
relational programming system. We prove in Coq that these two search strategies 
are are semantically equivalent, since \texttt{(run n ? g)} produces the same 
result in both our improved \BFS{} and BFS. (See supplements for the formal proof.) To compare efficiency, we 
translate \BFS{}'s Haskell code into Racket (See supplements for the 
translated code). The translation is direct due to the similarity of the two 
relational programming systems. The translated code is longer than our \BFS{}. 
And it runs slower in all benchmarks. Details about differences in efficiency 
are in \autoref{compare-efficiency}.

\section{quantitative evaluation}

>>>>

\usepackage{listings}
\lstset{
  language=Scheme,
  basicstyle=\ttfamily,
  morekeywords={run,conde,run*,defrel,==,fresh,disj,conj,disj+},
  alsodigit={!\$\%&*+-./:<=>?@^_~},
  morecomment=[l]{;,\#{}lang},
  mathescape=true,
  literate={-inf}{{$^\infty$}}{1} {/fair}{{$_{fair}$}}{3}
  		   {→}{{$\rightarrow$}}{1}
  		   {×}{{$\times$}}{1}
  		   {==}{{$\equiv$}}{1}
  		   {conde}{{\textcolor[rgb]{0,.3,.7}{\textbf{cond$^e$}}}}{4}
           {appendo}{{append$^o$}}{6}
  		   {repeato}{{repeat$^o$}}{6}
  		   {disj2}{{disj$_2$}}{4}
  		   {conj2}{{conj$_2$}}{4}
  		   {g0}{{g$_0$}}{2}
  		   {g1}{{g$_1$}}{2}
  		   {g2}{{g$_2$}}{2}
  		   {gl}{{g$_l$}}{2}
  		   {gr}{{g$_r$}}{2}
  		   {DFSi }{{DFS$_{i}$}}{3}
  		   {DFSbi}{{DFS$_{bi}$}}{3}
  		   {DFSf }{{DFS$_{f}$}}{3}
}
%% scheme-list .

@@

\newcommand{\DFSi }[0]{DFS$_{i}$}
\newcommand{\DFSf }[0]{DFS$_{f}$}
\newcommand{\DFSbi}[0]{DFS$_{bi}$}
\newcommand{\BFS}[0]{BFS}
\newcommand{\BFSser}[0]{BFS}
\newcommand{\BFSimp}[0]{Our BFS}

@@

miniKanren is a family of relational programming languages.
\citet{Friedman:2005:RS:1121583,friedman_reasoned_2018} introduce
miniKanren and its implementation in \emph{The Reasoned Schemer} 
and \emph{The Reasoned Schemer, 2nd Ed} (TRS2). \citet{byrd2017unified} have 
demonstrated that miniKanren programs are useful in solving several difficult
problems. miniKanren.org contains the seeds of many difficult problems and 
their solutions.

\NOTE[DPF]{...  Mostly the point is to also include a reference to
the two dimitris and to the first OCamren (OCanren), whatever. It is up to you
in this paragraph.  ...}
\NOTE[LKC]{I have fixed the cite to TRS and TRS2.}
\NOTE[LKC]{What is the point of citing dmitris?}
\NOTE[LKC]{Are we citing OCanren to show that miniKanren is widely used? If 
yes, do we need to cite other implementations as well?}

\NOTE[DPF]{I have decided to agree with you that we should not use 
BFSser, so I am going to remove all occurrences.  
Also, we should say, ``Our implementation of BFS''
and then say, shortened to ``Our BFS'' and then you can use it in the table,
so one column should be Our BFS and the other should be BFS.}
\NOTE[DPF]{Occurrences of BFS (without Our) mean Seres et al. (We need only say
that once, or very occasionally}
\NOTE[DPF]{Is the spelling of reverso truly written without the last e?
Shouldn't it be reverseo with the appropriate raising of o.}
\NOTE[LKC]{Yes, it is reverso (not reverseo) in Dmitris' paper}

@@

@@

\NOTE[DPF]{Would the answer be the same if the list had been this ((a) (b))
or did you mean structurally the same?}
\NOTE[LKC]{I mean for all answer in the answer list, the answer's inner lists 
are pairwise \texttt{equal?}. It is impossible to see a \texttt{((a) (b))} in 
the answer list.}

@@

In this section, we review the implementation of interleaving depth-first 
search (\DFSi). We focus on parts that are relevant to this paper. TRS2,
chapter 10 and the appendix, ``Connecting the wires'', 
provides a comprehensive description of the 
miniKanren 
implementation but limited to unification constraints ($\equiv$).
Fig.~\ref{DFSi-0} and Fig.~\ref{DFSi-1} show parts that are later compared 
with other search 
strategies. We follow some conventions to name variables: \texttt{s}s name 
states; \texttt{g}s (possibly with subscript) name goals; 
variables ending with $^\infty$ name search spaces. Fig.~\ref{DFSi-0} shows the 
implementation of \disj. The 
first function, \disjtwo, implements binary disjunction. It applies the 
two disjunctive goals to the input state \texttt{s} and composes the two 
resulting search spaces with \texttt{append$^\infty$}. The following syntax 
definitions say \disj{} is right-associative. Fig.~\ref{DFSi-1} 
shows the implementation of \conj. The first function, \conjtwo{}, 
implements binary conjunction. 
It applies the \emph{first} goal to the input state, then applies the second 
goal to states in the resulting search space. The helper function 
\texttt{append-map$^\infty$} applies its input goal to states 
in its input search spaces and composes the resulting search spaces. It reuses 
\texttt{append$^\infty$} for search space composition. The following syntax 
definitions say \conj{} is also right-associative.

@@

\NOTE[DPF]{I am not in love with the two calls to null?. I can think of at least
two ways to fix it.  Pick one.}
\NOTE[LKC]{fixed}

@@

\NOTE[DPF]{The use of the references to Fig 7 twice tells me that something
has to change for the sake of clarity.}
\NOTE[LKC]{I guess you are talking about Table 1 becasue there is no Fig 7. I 
have change the first reference.}

@@

\NOTE[DPF]{If we forget to use {} after \reverso, the o ends up too near
the next word.  I fixed reverso on the second line of section 7, but there 
may be others.}
\NOTE[LKC]{I have checked all uses of similar macros by text search.}

@@

In this subsection, we compare the pure subset of our BFS with \BFS{}. We 
focus on the pure subset because \BFS{} is designed for a pure 
relational programming system. We prove in Coq that these two search strategies 
are are semantically equivalent, since \texttt{(run n ? g)} produces the same 
result in both our improved \BFS{} and BFS. (See supplements for the formal proof.) To compare efficiency, we 
translate \BFS{}'s Haskell code into Racket (See supplements for the 
translated code). The translation is direct due to the similarity of the two 
relational programming systems. The translated code is longer than our \BFS{}. 
And it runs slower in all benchmarks. Details about differences in efficiency 
are in \autoref{sec:quan}.
% \autoref{compare-efficiency}

\NOTE[LKC]{I ref the next section instead of the table}

\section{quantitative evaluation}
\label{sec:quan}
